---
title: "research_shell_ShinaMin"
author: "Shina Min"
date: "2025-01-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```


### Investigating the Gene-Cohort Interaction Between FTO and BMI
  Understanding the genetic and environmental determinants of **obesity** is crucial for effective public health interventions. This research explores **the FTO gene**, a key genetic determinant of body mass index (BMI), to assess whether its influence differs across birth cohorts. It examines whether individuals born after World War II (WWII), who were exposed to a higher-calorie diet, show a stronger genetic **effect of FTO on BMI** compared to those born before WWII. To manage large-scale multiomic datasets, this study employs **Linux-based tools** and high-performance **cluster computing** for data preprocessing and analysis.



### 1. Study Design and Data Collection
  This study leverages **multiomic datasets** from 5,000 individuals across 500 families in the United States and Denmark. These datasets include **genomic, epigenomic, transcriptomic, proteomic, metabolomic, and lipidomic profiles.** Families have been followed for over a decade, allowing the integration of **whole-genome sequencing and lipid profiling** to examine how FTO variants influence lipid metabolism and BMI. Given the dataset's complexity, cluster-based parallel computing is used for preprocessing and efficient data management.
 

```{r, echo=FALSE}
#table for comparing two cohorts
cohort_data = data.frame(
  Cohort_Group = c("Born Before WWII (≤1945)", "Born After WWII (>1945)"),
  Key_Characteristics = c("Limited food availability, lower calorie intake",
                          "Industrialized food systems, higher calorie intake"))

kable(cohort_data, caption = "Comparison of Pre-WWII and Post-WWII Cohorts")
```

### 2. Computational Analysis
Principal component analysis (PCA) is applied to lipidomic data to identify biomarkers associated with obesity. Longitudinal statistical modeling assesses whether FTO has a stronger effect on BMI in post-WWII cohorts. Linux shell scripting and cluster-based processing ensure computational efficiency for handling large datasets.

### Conclusion
This study clarifies the **gene-environment interaction** between **FTO and BMI**, providing insights into obesity risk. By using high-performance computing, it supports **precision medicine strategies** for targeted interventions.


